Parsing Tweets into Universal Dependencies

### 2018-04-15 paper.tex
section 2: annotation (tweaks and questions)

%\newcommand{\yjcomment}[1]{\textcolor{orange}{[$_\mathrm{L}^\mathrm{Y}$#1]}}
%\newcommand{\nascomment}[1]{\textcolor{blue}{[#1 ---\textsc{nas}]}}
%\newcommand{\yicomment}[1]{\textcolor{gray}{[#1 ---\textsc{yi}]}}
%\newcommand{\nss}[1]{\textcolor{magenta}{[$_\mathrm{S}^\mathrm{NS}$#1]}}

@@

In the different direction of splitting tokens, UD guidelines also suggest to merge
{\it multi-token words} (e.g., {\it 20 000}) into one single token in some special
cases. We witnessed a small number of tweets that contain multi-token words
(e.g., {\it Y O}, and {\it R E T W E E T}) but didn't combine them for simplification.
Such tokens only account for 0.07\% and we use the UD
  {\it goeswith} relation to resolve these cases in the dependency annotations.

@@

Before turning to UD annotations, we (re)annotated the data with 
POS tags, for consistency with other UD efforts,
which adopt the universal POS tagset.\footnote{A revised and extended version of \citet{PETROV12.274} with 17 tags.}
% The POS annotation of a word in tweets is generally based on its syntactic distribution, which indicates that the POS of word changes on different context.
% Sticking to the syntactic distribution makes it possible that a `$\heart$' symbol is tagged as a verb as shown in Figure \ref{fig:non-syn-toks}. \yjcomment{Need a better explanation.}
%We follow the UD V2 morphology guideline and use the Universal POS tags~ to tag the tweet tokens. For twitter specific linguistic phenomena, we have set different strategies and try to align with the UD conventions and UD\textunderscore English as close as possible.\footnote{\url{https://github.com/UniversalDependencies/UD\_English}}
In some cases,  non-corresponding tag conflicts arose between the UD English 
treebank conventions \cite[UD\_English;][]{Marneffe2014UniversalSD}\footnote{\url{https://github.com/UniversalDependencies/UD_English-EWT}}
and the conventions of \newcite{gimpel-EtAl:2011:ACL-HLT2011}.  %\yicomment{The conflicts
                                %should refer to a {\bf
                                %noncorresponding tag} from that of
                                %gimpel to UDEnglish, cuz they dont
                                %have the same tagset.} 
In these cases, we always
conformed to UD, enabling consistency (e.g., when we exploit the
existing UD\_English treebank in our parser for tweets, \S\ref{sec:parsing}).  For example,  
the nominal URL in Figure \ref{fig:informal-toks} is tagged as {\it
  other} ({\tt X}) and {\tt +} is tagged as {\it symbol} ({\tt SYM})
rather than {\it conjunction} ({\tt CCONJ}).  
% Such violation is resulted by our attempt of conforming the annotation to the {\it universal dependencies - English dependency treebank} (UD\_English), in which all the URLs are tagged as {\it other} and symbolic conjunctors are tagged as {\it symbol}.
% By conforming our annotation to the UD\_English data, hopefully, the tools we built on tweets can benefit from the dataset on broader domain. \yjcomment{need wording}

@@

Tokens that abbreviate multiple words, such as \emph{idc} are resolved to the POS of the syntactic head of the
expression, following UD conventions (in this example, the head \emph{care} is
a verb, so \emph{idc} is tagged as a verb).
% The informal syntactic tokens can generally be handled by their distribution.
% However, one of the exception is the {\tt phrasal abbreviations} in which one token carries multiple syntactic functions.
% In dealing with such phrases, we propose a routine that first recover their original forms, then, use the POS of their syntactic head word as the POS of the whole phrase.
% Therefore, {\it idc} (I don't care) is a verb because {\it care} is its syntactic head.
% Such routine works for most of the phrasal abbreviations.
When the token is not phrasal, we use the POS of the left-most
sub-phrase.  For example, \emph{mfw} (``my face when'') is tagged as a
noun (for \emph{face}).  % \footnote{Such failure can be resulted by the mis-match of phrase boundaries between the original form and the definition of `phrase' in UD. For example, `mfw' (my face when) doesn't make a phrase because `when' belongs to the subordination that modifies `my face'.}

@@

It is important to note that these types may, in some
contexts, have syntactic functions.
For example, besides being a discourse marker, \emph{RT} can
abbreviate the verb {\it retweet}, and emoticons and hashtags may be
used as content words within a sentence, and at-mentions can be normal vocative proper nouns; see Figure
\ref{fig:informal-toks}.  Therefore, the criteria for annotating a
token as non-syntactic must be context-dependent.
%To our knowledge, this issue has not previously been addressed in UD
%annotation efforts \nascomment{check!}\yjcomment{it's not safe to claim that.}.

@@

\begin{itemize}
%\item For the at-mentions in retweet constructions, we attach them to their retweet markers. We will discuss our choice in the following paragraph;
\item If a non-syntactic token is within a sentence that has a clear predicate, it will be attached to this predicate. Retweet construction is a special case and we will discuss our choice in the following paragraph.
\item If the whole sentence is made of a sequence of non-syntactic tokens, we attach all these tokens to the first one.
%\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item Non-syntactic tokens are mostly labeled as {\it discourse}, but
  URLs are always labeled as {\it list}, following the UD\_English-EWT dataset.% \yicomment{need to add punct for punctuations?}
%\\ \yicomment{For retweet at-mention case, we attach the retweet at-mentions to the RT sign, so that the retweet action and the people whose tweet is being retweeted could be connected.\footnote{We can also follow the first convention to attach the retweet at-mention to its predicate, but we think it is better for downstream tasks.}}
\end{itemize}

@@

\newcite{kong-EtAl:2014:EMNLP2014} proposed an additional
preprocessing step, \emph{token selection}, in their annotation process.
They required the annotators to first select the non-syntactic
tokens and exclude them from the final dependencies annotation.
%  \nascomment{confusing.  were they removed or
%  just labeled as non-syntactic and then excluded from the tree?  I
%  think they were not actually removed, and were still available,
%  e.g., for defining features of the input for parsing}
In order to keep our annotation conventions in line with UD norms and preserve the original tweets as much as possible,
we include non-syntactic tokens in our annotation following the
conventions above. Compared with \citet{kong-EtAl:2014:EMNLP2014},
we also gave a clear %\yicomment{much clearer}
definition of non-syntactic tokens, which helped
us avoid confusion during annotation.

@@

\subsection{Comparison to PoSTWITA-UD}\label{sec:postwita}
The first Twitter treebank annotated with Universal Dependencies 
was the \mbox{\textit{PosTWITA-UD}} corpus for Italian \citep{sanguinetti-17}, 
which consists of 6,738 tweets (119,726 tokens).
In their convention, tokenization tends to preserve the original
tweet content but two special cases, \textit{articulated prepositions}
(e.g.~\textit{nella} as \textit{in la})
and \textit{clitic clusters} (e.g.~\textit{guardandosi} as \textit{guardando si}), are tokenized.
%\nss{explain what these two phenomena are. are these the only words that are split?}\yjcomment{done}
%Beyond tokenization, an extra layer of lemmas is annotated
%to recover ``what is understandable by a human''.
%It's arguable whether normalization (the lemmatization in \citet{sanguinetti-17})
%on social media data is essential according to \citet{eisenstein:2013:NAACL-HLT}.
%In this paper, we don't include normalization and leave it an open question for further study.
%\nss{Maybe just say: 
Their lemmas include spelling normalization, whereas our lemmas only normalize casing and inflectional morphology.
The current UD guidelines on lemmas give a lot of flexibility on treebank development.\footnote{\url{http://universaldependencies.org/u/overview/morphology.html\#lemmas}}
%You could also mention that the current UD guidelines on lemmas (which I wrote) give a lot of flexibility to treebank 
%developers: \url{http://universaldependencies.org/u/overview/morphology.html\#lemmas}}

%In the syntax annotation, tweet specific
%constructions are considered by both \citet{sanguinetti-17} and our paper.
With respect to tweet-specific constructions,
\citeposs{sanguinetti-17} and our  interpretation of headedness is the same, 
but we differ in the relation label.
For topical hashtags, we use \textit{discourse} while \citet{sanguinetti-17} used \textit{parataxis}. 
In referential URLs, we use \textit{list} (following the precedent of UD\_English-EWT) while \citet{sanguinetti-17} used \textit{dep}.
%<<<<<<< HEAD
%The choice of relation on the non-syntactic tokens is questionable
%since UD doesn't specify a convention for this. 
%Our choice of \textit{discourse} is inspired by the observation that emoticon is annotated as \textit{discourse}.
%\citet{sanguinetti-17} used the same relation for the emoticon but analyze topical hashtag as \textit{parataxis}.%\nss{what did PosTWITA-UD use for non-syntactic tokens?}\yjcomment{done}
%=======
%The choice of relation on the non-syntactic tokens is questionable
%since UD doesn't specify a convention for this. 
Our choice of \textit{discourse} for sentiment emoticons
is inspired by the observation that emoticons are annotated as \textit{discourse} by UD\_English-EWT
and \citet{sanguinetti-17} used the same relation for the emoticons.
%\nss{topical hashtags already mentioned above}
%>>>>>>> a7f86b777a73f64b2f05101f4804421aea0412f3
% Non-syntactic tokens can be easily recovered
% from our annotation and changing it to other relation in the future development is not difficult.
We expect that the official UD guidelines will eventually adopt standards for these constructions so the treebanks can be harmonized.

@@

\subsubsection{Data Collection}
{\sc Tweebank v2} is built on the original data of {\sc Tweebank v1}
(840 unique tweets, 639/201 for training/test set), along with an 
additional 210 tweets sampled from the POS-tagged dataset of
\newcite{gimpel-EtAl:2011:ACL-HLT2011} and 2,500 tweets sampled
from the Twitter stream from February 2016 to July 2016.\footnote{Data downloaded from \url{https://archive.org/}.}
The latter data source consists of 147.4M English tweets after being
filtered by the {\it lang} attribute in the tweet JSON and {\it
  langid.py}.\footnote{\url{https://github.com/saffsd/langid.py}}
In the same way as \citet{kong-EtAl:2014:EMNLP2014},
reference unit is always the tweet in its entirety
-- which may thus consist of multiple sentences -- not the sentence alone.
Before annotation, we use simple regular expression to anonymize username and URL.

@@

%	\begin{tabular}{rrrrrr}
%		\hline
%		\multicolumn{2}{l}{version} & train & dev & test & total \\
%		\hline
%		\textsc{v1} & tweets & 639 & 0 & 201 & 840 \\
%		& tokens  & 9,310 & 0 & 2,839 & 12149 \\
%		\hdashline
%		\textsc{v2} & tweets & 1,639 & 709 & 1,201 & 3,549 \\
%		 & tokens &	24,753	&	11,742	&	19,112 & 55,617\\
%		\hline
%	\end{tabular}
\caption{Statistics of {\sc Tweebank v2} and the comparison with
{\sc Tweebank v1} in quantity.
%\nss{not obvious how to interpret this: train tokens looks like 1,609+1,000 = 2,609, but maybe it's 609+1,000=1,609? 
%I think it would be easier just to show v1 and v2 as separate rows. Also, a column for totals across the splits.}
%\yjcomment{i've fix this.}
\label{tbl:data-stat}
}
\end{table}
Our annotation process was conducted in two stages.
In the first stage, 18 researchers worked on the {\sc Tweebank v1}
proportion and the additional 210 tweets and created the initial annotations in one day.
Before annotating, they were given a tutorial overview of the general UD
annotation conventions and our guidelines specifically for annotating tweets.
Both the guidelines and annotations
were further refined by the authors of this paper to increase
the coverage of our guidelines and solve inconsistencies between
different annotators during this exercise. In the second stage, a tokenizer, a POS tagger, and a
parser were trained on the annotated data from the first stage (1,050 tweets in total),
and used to automatically analyze the sampled 2,500 tweets.  Authors 
of this paper manually corrected the parsed data and finally achieved 3,550 labeled tweets.\footnote{Manual annotation was done with Arborator \cite{gerdes:2013:W13-37}, a web platform for drawing dependency trees.}
Newly created annotations are split into train, development, and test set and appended
to the original split of {\sc Twebank v1}. Statistics of our annotations and data split are shown
in Table \ref{tbl:data-stat}.

@@

We report the inter-annotator agreement between the annotators
in the second stage. There is very little disagreement on the
tokenization annotation. % \nascomment{still better to quantify}.
The agreement on POS is 96.6\%, the unlabeled
dependency agreement is 88.8\% and the labeled dependency agreement
is 84.3\%. Further analysis shows the major disagreements on POS
involve entity names (30.6\%) %\yicomment{noun phrases or proper noun
                              %phrases, do we need example
                              %here?}\yjcomment{i wasn't counting
                              %PROPN and NOUN confusion, but any type
                              %of confusion related with PROPN}
                              %\nascomment{probably don't need examples}
and topical hashtags (18.1\%).
Taking the example in Figure \ref{fig:non-syn-toks}, ``Fix you'' 
%\yicomment{need to relate to previous error analysis, entity names
%maybe?}
can
be understood as a verbal phrase but also as the name of the Coldplay's
single and tagged as proper noun. 
%\yicomment{need to say the reason why dependency inter-annotator
%agreement is low? 1. POS disagreement lead to big difference in
%understanding the semantics. 2. Some special constructions like
%ellipsis lead to different view of tweets. Fig3 is an example of 2.}
%\nascomment{I don't think we have room to go deeper}
An example of a disagreement on
dependencies is shown in Figure \ref{fig:disagree}.  Depending on
whether this is an example of a zero copula construction, or a clause-modified
noun, either annotation is plausible.

>>>>

%\newcommand{\yjcomment}[1]{\textcolor{orange}{[$_\mathrm{L}^\mathrm{Y}$#1]}}
%\newcommand{\nascomment}[1]{\textcolor{blue}{[#1 ---\textsc{nas}]}}
%\newcommand{\yicomment}[1]{\textcolor{gray}{[#1 ---\textsc{yi}]}}
\newcommand{\nss}[1]{\textcolor{magenta}{[$_\mathrm{S}^\mathrm{NS}$#1]}}

@@

In the different direction of splitting tokens, UD guidelines also suggest to merge
{\it multi-token words} (e.g., {\it 20 000}) into one single token in some special
cases. We witnessed a small number of tweets that contain multi-token words
(e.g., {\it Y O}, and {\it R E T W E E T}) but didn't combine them for simplicity.
Such tokens only account for 0.07\% and we use the UD
  {\it goeswith} relation to resolve these cases in the dependency annotations.

@@

Before turning to UD annotations, we (re)annotated the data with 
POS tags, for consistency with other UD efforts,
which adopt the universal POS tagset.\footnote{A revised and extended version of \citet{PETROV12.274} with 17 tags.}
% The POS annotation of a word in tweets is generally based on its syntactic distribution, which indicates that the POS of word changes on different context.
% Sticking to the syntactic distribution makes it possible that a `$\heart$' symbol is tagged as a verb as shown in Figure \ref{fig:non-syn-toks}. \yjcomment{Need a better explanation.}
%We follow the UD V2 morphology guideline and use the Universal POS tags~ to tag the tweet tokens. For twitter specific linguistic phenomena, we have set different strategies and try to align with the UD conventions and UD\textunderscore English as close as possible.\footnote{\url{https://github.com/UniversalDependencies/UD\_English}}
In some cases,  non-corresponding tag conflicts arose between the UD English Web Treebank 
treebank conventions \cite[UD\_English-EWT;][]{Marneffe2014UniversalSD}\footnote{\url{https://github.com/UniversalDependencies/UD_English-EWT}}
and the conventions of \newcite{gimpel-EtAl:2011:ACL-HLT2011}.  %\yicomment{The conflicts
                                %should refer to a {\bf
                                %noncorresponding tag} from that of
                                %gimpel to UDEnglish, cuz they dont
                                %have the same tagset.} 
In these cases, we always
conformed to UD, enabling consistency (e.g., when we exploit the
existing UD\_English-EWT treebank in our parser for tweets, \S\ref{sec:parsing}).  For example,  
the nominal URL in Figure \ref{fig:informal-toks} is tagged as {\it
  other} ({\tt X}) and {\tt +} is tagged as {\it symbol} ({\tt SYM})
rather than {\it conjunction} ({\tt CCONJ}).  
% Such violation is resulted by our attempt of conforming the annotation to the {\it universal dependencies - English dependency treebank} (UD\_English), in which all the URLs are tagged as {\it other} and symbolic conjunctors are tagged as {\it symbol}.
% By conforming our annotation to the UD\_English data, hopefully, the tools we built on tweets can benefit from the dataset on broader domain. \yjcomment{need wording}

@@

Tokens that abbreviate multiple words (such as \emph{idc}) are resolved to the POS of the syntactic head of the
expression, following UD conventions (in this example, the head \emph{care} is
a verb, so \emph{idc} is tagged as a verb).
% The informal syntactic tokens can generally be handled by their distribution.
% However, one of the exception is the {\tt phrasal abbreviations} in which one token carries multiple syntactic functions.
% In dealing with such phrases, we propose a routine that first recover their original forms, then, use the POS of their syntactic head word as the POS of the whole phrase.
% Therefore, {\it idc} (I don't care) is a verb because {\it care} is its syntactic head.
% Such routine works for most of the phrasal abbreviations.
When the token is not phrasal, we use the POS of the left-most
sub-phrase.  For example, \emph{mfw} (``my face when'') is tagged as a
noun (for \emph{face}).  % \footnote{Such failure can be resulted by the mis-match of phrase boundaries between the original form and the definition of `phrase' in UD. For example, `mfw' (my face when) doesn't make a phrase because `when' belongs to the subordination that modifies `my face'.}

@@

It is important to note that these types may, in some
contexts, have syntactic functions.
For example, besides being a discourse marker, \emph{RT} can
abbreviate the verb {\it retweet}; emoticons and hashtags may be
used as content words within a sentence; and at-mentions can be normal vocative proper nouns: see Figure~\ref{fig:informal-toks}.  
Therefore, the criteria for annotating a
token as non-syntactic must be context-dependent.
%To our knowledge, this issue has not previously been addressed in UD
%annotation efforts \nascomment{check!}\yjcomment{it's not safe to claim that.}.

@@

\begin{itemize}
%\item For the at-mentions in retweet constructions, we attach them to their retweet markers. We will discuss our choice in the following paragraph;
\item If a non-syntactic token is within a sentence that has a clear predicate, it will be attached to this predicate. 
The retweet construction is a special case and we will discuss its treatment in the following paragraph.
\item If the whole sentence\nss{tweet?} is a sequence of non-syntactic tokens, we attach all these tokens to the first one.
%\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item Non-syntactic tokens are mostly labeled as {\it discourse}, but
  URLs are always labeled as {\it list}, following the UD\_English-EWT dataset.% \yicomment{need to add punct for punctuations?}
%\\ \yicomment{For retweet at-mention case, we attach the retweet at-mentions to the RT sign, so that the retweet action and the people whose tweet is being retweeted could be connected.\footnote{We can also follow the first convention to attach the retweet at-mention to its predicate, but we think it is better for downstream tasks.}}
\end{itemize}

@@

\newcite{kong-EtAl:2014:EMNLP2014} proposed an additional
preprocessing step, \emph{token selection}, in their annotation process.
They required the annotators to first select the non-syntactic
tokens and exclude them from the final dependency annotation.
%  \nascomment{confusing.  were they removed or
%  just labeled as non-syntactic and then excluded from the tree?  I
%  think they were not actually removed, and were still available,
%  e.g., for defining features of the input for parsing}
In order to keep our annotation conventions in line with UD norms and preserve the original tweets as much as possible,
we include non-syntactic tokens in our annotation following the
conventions above. Compared with \citet{kong-EtAl:2014:EMNLP2014},
we also gave a clear %\yicomment{much clearer}
definition of non-syntactic tokens, which helped
us avoid confusion during annotation.

@@

\subsection{Comparison to PoSTWITA-UD}\label{sec:postwita}
The first Twitter treebank annotated with Universal Dependencies 
was the \mbox{\textit{PosTWITA-UD}} corpus for Italian \citep{sanguinetti-17}, 
which consists of 6,738 tweets (119,726 tokens).
In their convention, tokenization tends to preserve the original
tweet content but two special cases, \textit{articulated prepositions}
(e.g.~\textit{nella} as \textit{in la})
and \textit{clitic clusters} (e.g.~\textit{guardandosi} as \textit{guardando si}), are tokenized.
%\nss{explain what these two phenomena are. are these the only words that are split?}\yjcomment{done}
%Beyond tokenization, an extra layer of lemmas is annotated
%to recover ``what is understandable by a human''.
%It's arguable whether normalization (the lemmatization in \citet{sanguinetti-17})
%on social media data is essential according to \citet{eisenstein:2013:NAACL-HLT}.
%In this paper, we don't include normalization and leave it an open question for further study.
%\nss{Maybe just say: 
Their lemmas include spelling normalization, whereas our lemmas only normalize casing and inflectional morphology.
The current UD guidelines on lemmas are flexible, so variation between treebanks is expected.\footnote{\url{http://universaldependencies.org/u/overview/morphology.html\#lemmas}}
%You could also mention that the current UD guidelines on lemmas (which I wrote) give a lot of flexibility to treebank 
%developers: \url{http://universaldependencies.org/u/overview/morphology.html\#lemmas}}

%In the syntax annotation, tweet specific
%constructions are considered by both \citet{sanguinetti-17} and our paper.
With respect to tweet-specific constructions,
\citeposs{sanguinetti-17} and our  interpretation of headedness is the same, 
but we differ in the relation label.
For topical hashtags, we use \textit{discourse} while \citet{sanguinetti-17} used \textit{parataxis}. 
In referential URLs, we use \textit{list} (following the precedent of UD\_English-EWT) while \citet{sanguinetti-17} used \textit{dep}.
%<<<<<<< HEAD
%The choice of relation on the non-syntactic tokens is questionable
%since UD doesn't specify a convention for this. 
%Our choice of \textit{discourse} is inspired by the observation that emoticon is annotated as \textit{discourse}.
%\citet{sanguinetti-17} used the same relation for the emoticon but analyze topical hashtag as \textit{parataxis}.%\nss{what did PosTWITA-UD use for non-syntactic tokens?}\yjcomment{done}
%=======
%The choice of relation on the non-syntactic tokens is questionable
%since UD doesn't specify a convention for this. 
Our choice of \textit{discourse} for sentiment emoticons
is inspired by the observation that emoticons are annotated as \textit{discourse} by UD\_English-EWT
and \citet{sanguinetti-17} used the same relation for the emoticons.\nss{how does the other treebank treat the RT construction and truncated words?}
%\nss{topical hashtags already mentioned above}
%>>>>>>> a7f86b777a73f64b2f05101f4804421aea0412f3
% Non-syntactic tokens can be easily recovered
% from our annotation and changing it to other relation in the future development is not difficult.
We expect that the official UD guidelines will eventually adopt standards for these constructions so the treebanks can be harmonized.

@@

\subsubsection{Data Collection}
{\sc Tweebank v2} is built on the original data of {\sc Tweebank v1}
(840 unique tweets, 639/201 for training/test), along with an 
additional 210 tweets sampled from the POS-tagged dataset of
\newcite{gimpel-EtAl:2011:ACL-HLT2011} and 2,500 tweets sampled
from the Twitter stream from February 2016 to July 2016.\footnote{Data downloaded from \url{https://archive.org/}.}
The latter data source consists of 147.4M English tweets after being
filtered by the {\it lang} attribute in the tweet JSON and {\it
  langid.py}.\footnote{\url{https://github.com/saffsd/langid.py}}
In the same way as \citet{kong-EtAl:2014:EMNLP2014},
reference unit is always the tweet in its entirety
-- which may thus consist of multiple sentences -- not the sentence alone.\nss{``Each tweet is annotated with a single tree, rather than being split into sentences.''?}
Before annotation, we use simple regular expression to anonymize usernames and URLs.

@@

%	\begin{tabular}{rrrrrr}
%		\hline
%		\multicolumn{2}{l}{version} & train & dev & test & total \\
%		\hline
%		\textsc{v1} & tweets & 639 & 0 & 201 & 840 \\
%		& tokens  & 9,310 & 0 & 2,839 & 12149 \\
%		\hdashline
%		\textsc{v2} & tweets & 1,639 & 709 & 1,201 & 3,549 \\
%		 & tokens &	24,753	&	11,742	&	19,112 & 55,617\\
%		\hline
%	\end{tabular}
\caption{Statistics of {\sc Tweebank v2} and the comparison with
{\sc Tweebank v1} in quantity.
%\nss{not obvious how to interpret this: train tokens looks like 1,609+1,000 = 2,609, but maybe it's 609+1,000=1,609? 
%I think it would be easier just to show v1 and v2 as separate rows. Also, a column for totals across the splits.}
%\yjcomment{i've fix this.}
\label{tbl:data-stat}
}
\end{table}
Our annotation process was conducted in two stages.
In the first stage, 18 researchers worked on the {\sc Tweebank v1}
portion and the additional 210 tweets and created the initial annotations in one day.
Before annotating, they were given a tutorial overview of the general UD
annotation conventions and our guidelines specifically for annotating tweets.
Both the guidelines and annotations
were further refined by the authors of this paper to increase
the coverage of our guidelines and solve inconsistencies between
different annotators during this exercise. In the second stage, a tokenizer, a POS tagger, and a
parser were trained on the annotated data from the first stage (1,050 tweets in total),
and used to automatically analyze the sampled 2,500 tweets.  Authors 
of this paper manually corrected the parsed data and finally achieved 3,550 labeled tweets.\footnote{Manual annotation was done with Arborator \cite{gerdes:2013:W13-37}, a web platform for drawing dependency trees.}
Newly created annotations are split into train, development, and test sets and appended
to the original splits of {\sc Twebank v1}. Statistics of our annotations and data splits are shown
in Table~\ref{tbl:data-stat}.

@@

We report the inter-annotator agreement between the annotators
in the second stage. There is very little disagreement on the
tokenization annotation. % \nascomment{still better to quantify}.
The agreement rate is 96.6\% on POS, 
88.8\% on unlabeled dependencies, and
84.3\% on labeled dependencies. 
Further analysis shows the major disagreements on POS
involve entity names (30.6\%) %\yicomment{noun phrases or proper noun
                              %phrases, do we need example
                              %here?}\yjcomment{i wasn't counting
                              %PROPN and NOUN confusion, but any type
                              %of confusion related with PROPN}
                              %\nascomment{probably don't need examples}
and topical hashtags (18.1\%).
Taking the example in Figure \ref{fig:non-syn-toks}, ``Fix you'' 
%\yicomment{need to relate to previous error analysis, entity names
%maybe?}
can
be understood as a verbal phrase but also as the name of the Coldplay's
single and tagged as proper noun. 
%\yicomment{need to say the reason why dependency inter-annotator
%agreement is low? 1. POS disagreement lead to big difference in
%understanding the semantics. 2. Some special constructions like
%ellipsis lead to different view of tweets. Fig3 is an example of 2.}
%\nascomment{I don't think we have room to go deeper}
An example of a disagreement on
dependencies is shown in Figure \ref{fig:disagree}.  Depending on
whether this is an example of a zero copula construction, or a clause-modified
noun, either annotation is plausible.
